### 2020-10-09 input/figdelA.tex
wip

\begin{tikzpicture}[scale=0.87]

  \newcommand\X{50pt}
  \newcommand\Y{-50pt}

  \newcommand\ADD{Add}
  \newcommand\DEL{Del}

  \draw (-\X, 0) -- node[above,font=\tiny]{$\ADD_{a, 1.5}$} (0, 0); %% a - b
  \draw (0, 0) -- node[above, font=\tiny]{$\ADD_{c, 1}$} (\X, 0); %% b - c
  \draw (0, 0) -- (0, \Y); %% b - d
  \draw (\X, 0) -- node[below, font=\tiny, sloped]{$\ADD_{c, 1}$} (0, \Y); %% c - d

  \draw[fill=white] (-\X, 0) node{$\bm{a}$} +(-5pt, -5pt) rectangle +(5pt, 5pt);  
  \draw[fill=white] (0, 0) node{$\bm{b}$} +(-5pt, -5pt) rectangle +(5pt, 5pt);
  \draw[fill=white] (\X, 0) node{$\bm{c}$} +(-5pt, -5pt) rectangle +(5pt, 5pt);
  \draw[fill=white] (0, \Y) node{$\bm{d}$} +(-5pt, -5pt) rectangle +(5pt, 5pt);

  \draw (-\X+5pt, 5pt) node[above left]{$\DEL_a$};
  \draw (\X+5pt, 5pt) node[above right]{$\DEL_c$};

\end{tikzpicture}