### 2021-02-15 input/figproofA.tex
figures for proofs

\begin{tikzpicture}[scale=0.87]

  \thickmuskip=0mu
  \medmuskip=0mu
  \thinmuskip=0mu

  \newcommand\X{50pt}
  \newcommand\Y{-45pt}

  \newcommand\ADD{\alpha}
  \newcommand\DEL{\delta}

  \draw[opacity=0] (-2*\X, 0) -- (3*\X, 0);

  \draw[->] (-\X + 5pt, 0) -- node[above, color=\PA, font=\small]{$\ADD_a^1$} (0 - 5pt, 0); %% a - b

  \draw[->] (0 +5pt, 0) -- (\X -5pt, 0); %% b - ...

  \draw[->] ( \X +5pt, 0) -- (2*\X -5pt, 0); %% ... - c

  \draw[->] ( 0, \Y + 5pt) -- ( 0, -5pt); %% ... - b

  \draw[->] ( -\X + 5pt, \Y ) -- ( -5pt, \Y); %% d - ...

  \draw[fill=white] (-\X, 0) node{$\bm{a}$} +(-5pt, -5pt) rectangle +(5pt, 5pt);  
  \draw[fill=white] (0, 0) node{$\bm{b}$} +(-5pt, -5pt) rectangle +(5pt, 5pt);
  \draw[fill=white] (\X, 0) node{$\ldots$} +(-5pt, -5pt) rectangle +(5pt, 5pt);
  \draw[fill=white] (2*\X, 0) node{$\bm{c}$} +(-5pt, -5pt) rectangle +(5pt, 5pt);

  \draw[fill=white] ( 0, \Y) node{$\ldots$} +(-5pt, -5pt) rectangle +(5pt, 5pt);
  \draw[fill=white] (-\X, \Y) node{$\bm{d}$} +(-5pt, -5pt) rectangle +(5pt, 5pt);

  \draw (-\X, 5pt) node[above, color=\PA]{$\ADD_a^0$};
  \draw ( 0, 5pt) node[above, color=\PC]{$\ADD_d^{0.5}$};
  \draw (2*\X, 5pt) node[above]{expect \textcolor{\PA}{$\alpha_a$}};

  \draw (-\X, \Y + 5pt) node[above, color=\PC]{$\DEL_d$};

\end{tikzpicture}