### 2021-02-26 input/figASmotivation.tex
mostly about AS

\begin{tikzpicture}[scale=1]

  \thickmuskip=0mu
  \medmuskip=0mu
  \thinmuskip=0mu

  \newcommand\X{50pt}
  \newcommand\Y{-50pt}

  \newcommand\ADD{\alpha}

  \draw (-1.5*\X, \Y + 5pt)
  -- node[above, font=\scriptsize, sloped]{$10ms$}
  (-1.5*\X, 0 - 5pt); %% a-e

  \draw (-1.5*\X + 5pt, 0)
  --
  (-1*\X, 0);

  \draw[dotted] (-1*\X, 0) --
  node[above, font=\scriptsize]{$300ms$} (-0.5*\X, 0);

  \draw (-0.5*\X, 0)
  --  
  (0 - 5pt, 0); %% a - b

  \draw (0 +5pt, 0)
  -- node[above, font=\scriptsize, sloped]{$15ms$}  
  (\X -5pt, 0); %% b - c

  \draw (0, \Y + 5pt)
  -- node[above, font=\scriptsize, sloped]{$20ms$}  
  (0, 0 - 5pt); %% b - d

  \draw (\X + 3pt, 0 - 5pt)
  -- node[above, font=\scriptsize, sloped]{$10ms$}  
  (0 + 5pt, \Y - 3pt); %% c - d

  \draw[dashed] (-0.75*\X, -0.5*\Y) -- (-0.75*\X, 1.5*\Y)
  node[above left]{\textit{(FR) Paris -- AS 1 \vphantom{g}}}
  node[above right]{\textit{AS 2 -- Wellington (NZ)}};

  \draw[fill=white] (-1.5*\X, 0) node[color=\PA]{$\bm{a}$} +(-5pt, -5pt) rectangle +(5pt, 5pt);
  \draw [left] (-1.5*\X - 5pt, 0) node{\textbf{e: 1}};
  \draw[fill=white] (0, 0) node[color=\PA]{$\bm{b}$} +(-5pt, -5pt) rectangle +(5pt, 5pt);
  \draw [above] ( 0, 5pt ) node{\textbf{e: 2}};
  \draw[fill=white] (\X, 0) node{$\bm{c}$} +(-5pt, -5pt) rectangle +(5pt, 5pt);
  \draw[fill=white] (0, \Y) node{$\bm{d}$} +(-5pt, -5pt) rectangle +(5pt, 5pt);

  \draw[fill=white] (-1.5*\X, \Y) node[color=\PA]{$\bm{e}$} +(-5pt, -5pt) rectangle +(5pt, 5pt);
  \draw [left] (-1.5*\X - 5pt, \Y) node{\textbf{e: 0}};

  % \draw (-\X, 5pt) node[above, color=\PA]{$\ADD_a^0$};
  % \draw (\X, 5pt) node[above, color=\PC]{$\ADD_c^0$};

\end{tikzpicture}