### 2021-03-23 input/figproof.tex
reworkd figures of proofs

\begin{figure*}
  \begin{center}
    \subfloat[Part A][\label{fig:proofA}Stale $\alpha$'s may stop
      other $\alpha$'s from reaching all processes ($b$ and $c$) that
      require it along the shortest path. Hence, we must guarantee
      eventual purging of stale $\alpha$'s (see
      Figure~\ref{fig:proofB}).]  {\input{input/figproofA.tex}}
    \hspace{10pt}
    \subfloat[Part B][\label{fig:proofB}Stale $\alpha$'s may
      stop $\delta$'s from reaching
      processes with targeted $\alpha$'s.
      To ensure consistent partitioning, $b'$ and $b''$
      must remove $\alpha_{d}$ either by delivering $\delta_d$,
      or a more specific $\delta_d^{0.5}$ and $\delta_d^{1.5}$ respectively.]
    {\input{input/figproofB.tex}}
    \caption{\label{fig:proof}Dynamic partitioning leads to correctness issues due to
      staleness and ordering of operations.}
  \end{center}
\end{figure*}