### 2021-04-07 input/figproblem.tex
figures

\begin{figure*}
  \begin{center}
    \subfloat[Part A][\label{fig:problemA}Both Process $a$ and
      Process $c$ initiate then remove a partition.
      $w_{ab} = 1.5$; $w_{bc} = w_{cd} = w_{bd} = 1$.]{\input{input/figproblemA.tex}}
    \hspace{5pt}
    \subfloat[Part B][\label{fig:problemB}Process $b$ and Process $d$ receive and deliver
      messages based on their weights. $a$'s partition propagates to all processes.]
             {\input{input/figproblemB.tex}}
    \hspace{5pt}
    \subfloat[Part C][\label{fig:problemC}Process $c$'s partition reach $b$ and $d$.
      Its removal shortly follows. Process $b$ does not have any control
      information about partitions any more.]{\input{input/figproblemC.tex}}
    \hspace{5pt}
    \subfloat[Part D][\label{fig:problemD}Processes discard remaining messages.
      Process $b$ does not deliver $\delta_a$, for it already removed $a$ when
      Partition $c$ dominated it. Processes $c$ and $d$ remain in an incorrect
      partition.]{\input{input/figproblemD.tex}}
  \end{center}
  \caption{\label{fig:problem}Partition removals cannot rely only on
    current partition, for it may lead to either inconsistent
    partitioning, or every process receiving every broadcast.}
\end{figure*}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "../paper"
%%% ispell-local-dictionary: "english"
%%% End: