### 2021-04-09 input/figproblem.tex
still on examples mostly

\begin{figure*}
  \begin{center}
    \subfloat[Part A][\label{fig:problemA}Both Process~$a$ and
      Process~$d$ add then delete their partition. For the recall,
      $w_{ab} = 2$; $w_{bc} = w_{bd} = 1$; $w_{cd} =
      3$.]{\input{input/figproblemA.tex}}
    \hspace{5pt} \subfloat[Part B][\label{fig:problemB}Process~$b$
      delivers and forwards the $\alpha$ message about $P_a$ while
      Process~$c$ delivers and forwards both $\alpha$ and $\delta$
      messages about $P_d$.]  {\input{input/figproblemB.tex}}
    \hspace{5pt} \subfloat[Part C][\label{fig:problemC}Process~$b$
      already blocks the propagation of $\delta_a$ despite it
      propagated $\alpha_a$ before. Process~$c$ and Process~$d$ do not
      deliver $\alpha_d$, for they already deleted
      it before.]{\input{input/figproblemC.tex}}
    \hspace{5pt} \subfloat[Part D][\label{fig:problemD}Processes
      discard remaining messages. Since Process~$b$ cannot deliver
      $\delta_a$, $c$ and $d$ remain in the deleted partition
      $P_a$.]{\input{input/figproblemD.tex}}
  \end{center}
  \caption{\label{fig:problem}Naive propagation of $\alpha$ and
    $\delta$ messages is insufficient to guarantee consistent partitioning.}
\end{figure*}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "../paper"
%%% ispell-local-dictionary: "english"
%%% End: