### 2021-04-12 input/figproofA.tex

\begin{tikzpicture}[scale=0.87]

  \thickmuskip=0mu
  \medmuskip=0mu
  \thinmuskip=0mu

  \newcommand\X{85pt}
  \newcommand\Y{-40pt}

  \newcommand\ADD{\alpha}
  \newcommand\DEL{\delta}

  \draw[opacity=0] (-1.4*\X, 0) -- (1.4*\X, 0);

  \draw[<->] (-5pt - \X, -\Y) -- node[above]{I: propagation} (5pt, -\Y); %% I

  \draw[->] (-\X + 5pt, 0) --
  node[above, color=\PA, font=\small]{$\ADD_a^1$}
  (0 - 5pt, 0); %% a - b

  \draw[<->] (-5pt, 5pt -\Y) -- node[above]{II: need I', III, IV} (5pt + \X, 5pt -\Y); %% II
  \draw[->] (0 + 5pt, 0) -- (-5pt +  \X, 0); %% b - c

  \draw[fill=white] (-\X, 0) node{$\bm{a}$} +(-5pt, -5pt) rectangle +(5pt, 5pt);  
  \draw[fill=white] (0, 0) node{$\bm{b}$} +(-5pt, -5pt) rectangle +(5pt, 5pt);
  \draw[fill=white] (\X, 0) node{$\bm{c}$} +(-5pt, -5pt) rectangle +(5pt, 5pt);

  \draw (-\X, 5pt) node[above, color=\PA]{$\ADD_a^0$};
  \draw ( 0, 5pt) node[above, color=\WRONG]{$\ADD_d^{0.5}$};
  \draw (\X,-5pt) node[below, font=\scriptsize]{expect \textcolor{\PA}{$\alpha_a$}};
  \draw (0, -5pt) node[below, font=\scriptsize]{expect $\DEL_d \rightarrow \textcolor{\PA}{\alpha_a}$};

\end{tikzpicture}