### 2021-04-19 input/figproofB.tex

\begin{tikzpicture}[scale=0.87]

  \thickmuskip=0mu
  \medmuskip=0mu
  \thinmuskip=0mu

  \newcommand\X{85pt}
  \newcommand\Y{-50pt}

  \newcommand\ADD{\alpha}
  \newcommand\DEL{\delta}

  \draw[opacity=0] (-1.4*\X, 0) -- (2.4*\X, 0);

  \draw[->] (-1.15*\X, 0) -- (-5pt + -1*\X, 0); \draw[dotted] (-1.25*\X, 0) -- (-1.15*\X, 0);
  \draw (5pt + 2*\X, 0) -- (2.15*\X, 0); \draw[dotted,->] (2.15*\X, 0) -- (2.25*\X, 0);

  \draw[<->] (5pt - \X, 0.4*\Y) --
  node[below, font=\small]{\textbf{I': propagation}}
  (-5pt, 0.4*\Y); %% I as well

  \draw[->] (-\X + 5pt, 0)
  node[below=-0.3em, below right, font=\scriptsize]{$\DEL_d \rightarrow$}
  -- (0 - 5pt, 0)
  node [below=-0.15em, below left, font=\tiny]{$\textcolor{\WRONG}{d \neq f}$}
  ; %% d - e

  \draw[<->] (5pt , 0.4*\Y) --
  node[below, font=\small]{\textbf{III: detection}}
  (-5pt + \X, 0.4*\Y); %% IV

  \draw[->] (0 +5pt, 0)
  node[below=-0.3em, below right, font=\tiny]{$\textcolor{\PF}{\ADD_f^{z'}} \rightarrow$}
  --
  (\X -5pt, 0)
  node [left= -0.15em, below=-0.3em, below left, font=\tiny]{$\textcolor{\WRONG}{\DEL_f \not\rightarrow \ADD_f^{z'}}$}
  ; %% e - b'

  \draw[<->] (5pt +\X , 0.4*\Y) --
  node[below, font=\small]{\textbf{IV: propagation}}
  (-5pt + 2*\X, 0.4*\Y); %% V

  \draw[->] ( \X +5pt, 0) -- (2*\X -5pt, 0); %% b - b''

  \draw[fill=white] (-\X, 0) node{$\bm{d}$} +(-5pt, -5pt) rectangle +(5pt, 5pt);  
  \draw[fill=white] (0, 0) node[color=\PF]{$\bm{e}$} +(-5pt, -5pt) rectangle +(5pt, 5pt);
  \draw[fill=white] (\X, 0) node[color=\PD]{$\bm{b'}$} +(-5pt, -5pt) rectangle +(5pt, 5pt);
  \draw[fill=white] (2*\X, 0) node[color=\PD]{$\bm{b''}$} +(-5pt, -5pt) rectangle +(5pt, 5pt);

  \draw (-\X, 5pt) node[above]{$\textcolor{\PD}{\ADD_d} \rightarrow \DEL_d$};

  \draw ( 0, 5pt) node[above, font=\footnotesize]{$
    \xrightarrow[\textcolor{\PC}{\ADD_d^{\vphantom{z'}}}]{\mbox{\scriptsize{before}}}
    \xrightarrow[\textcolor{\PF}{\ADD_f^{z\vphantom{'}}}]{\mbox{\scriptsize{last}}}
    \xrightarrow[\DEL_f^{\vphantom{z'}} ]{\mbox{\scriptsize{expect}}}$};

  \draw ( \X, 5pt) node[above, font=\footnotesize]{$
    \xrightarrow[\textcolor{\PF}{\ADD_f^{\vphantom{y'}}} \rightarrow \DEL_f]{\mbox{\scriptsize{before}}}
    \xrightarrow[\textcolor{\PC}{\ADD_d^{y'}}]{\mbox{\scriptsize{last}}}
    \xrightarrow[\DEL_d^{\vphantom{z'}}?]{\mbox{\scriptsize{expect}}}$};

  \draw (2*\X, 5pt) node[above, font=\footnotesize]{$
    \xrightarrow[\textcolor{\PC}{\ADD_d^{y''}}]{\mbox{\scriptsize{last}}}
    \xrightarrow[\DEL_d^{\vphantom{y'}}?]{\mbox{\scriptsize{expect}}}$};

  \draw[->] (-\X,  0.5*\Y) -- (-\X, -5pt);
  \draw[dotted] (-\X,  0.5*\Y) -- (-\X, 0.9*\Y);

  \draw[->] ( 0,  0.5*\Y) -- ( 0, -5pt);
  \draw[dotted] ( 0,  0.5*\Y) -- ( 0, 0.9*\Y) node[below right, font=\scriptsize]{somewhere: $\textcolor{\PF}{\ADD_f} \rightarrow \DEL_f$};

  \draw[->] ( \X,  0.5*\Y) -- ( \X, -5pt);
  \draw[dotted] ( \X,  0.5*\Y) -- ( \X, 0.9*\Y);

  \draw[->] (2*\X,  0.5*\Y) -- (2*\X, -5pt);
  \draw[dotted] (2*\X,  0.5*\Y) -- (2*\X, 0.9*\Y);

\end{tikzpicture}