### 2021-04-23 input/related_work.tex

\section{Related Work}
\label{sec:related_work}

Content indexing in geo-distributed infrastructures has been studied
through various approaches.  Offering an indexing service which is
either hosted by a centralized entity or distributed through nodes of
the network has long been the predominant one. This mainly requires
(i) to have access to the list of current replica locations and (ii)
to be aware of the topology to compute which replica is the closest
for a given location.

(i) Deploying a centralized server that maintains an index of all
replica locations has been proposed in~\cite{snamp, p2p-oracle,
  fogstore, p2p-alto}. These solutions suffer from well-known
disadvantages inherent to centralized solutions (load balancing,
robustness, locality, etc.) that are even magnified in an Edge
context.  Distributing this index among nodes in the network
circumvent some of these issues, either using coordinate-based
overlays~\cite{voronet, coin_19} or \underline{D}istributed
\underline{H}ash \underline{T}ables (DHT)~\cite{ipfs, mdht, squirrel}
for example. In the latter, each node stores a part of the index,
defined by an interval between hash values.  Before downloading any
object, a node first hashes the content to obtain the address of the
node that stores the replica locations of this given object. After
obtaining from the remote node the list of available replicas, it can
select the closest copy to download from. We emphasize that, in this
setting, the DHT only stores information about replicas locations, not
the content itself. DHT-based approaches are well-studied and easy to
deploy in order to offer such a localization service.

(ii) Being centralized or distributed, this index should be enhanced
with the topology knowledge in order to be able to decide where
resides the closest replica. Maintaining a consistent view of an ever
changing topology across a network is inherently complicated,
especially in asynchronous settings. This is a well-studied problem in
the network community and numerous protocols have been proposed. As
for the index, this knowledge about the topology can be either
centralized~\cite{topology-discovery} or distributed like in routing
protocols such as \underline{O}pen \underline{S}hortest
\underline{P}ath \underline{F}irst (OSPF).

All these solutions require to contact a remote node (either a server
or a DHT node) to request this index which is costly in terms of
latency while in contradiction with Edge infrastructure constraints,
as previously explained.

The opposite approach consists in hosting the index directly on the
nodes, avoiding any prior request to access a content.
%This requires notifying nodes of replica creations and removals along
%with information about the evolving topology.
Broadcasting information about cache updates to all nodes in the
system is a straightforward way to maintain consistent information
about replica locations for all nodes, as usually implemented in
\underline{N}amed \underline{D}ata \underline{N}etworking
(NDN)~\cite{nlsr}.  Indeed, having the entire knowledge of all the
replica localization along with distance information carried into
messages, each node can easily determine where the closest copy of a
given object resides, without contacting any remote node.  Similarly,
one could solve this issue by using a
\underline{C}onflict-\underline{F}ree \underline{R}eplicated
\underline{D}atatype (CRDT) for set data structures, for
example~\cite{shapiro2011crdts}. Yet both solutions inherently imply
that every node will eventually receive all messages.

At first glance, membership protocols as proposed in~\cite{t-man}
would appear to provide suitable abstractions to create a
scope-related partitioning, as proposed in this paper. Such protocols
usually rely on gossip mechanism, implying contacts with random nodes
that can be very far away in the physical topology. Moreover,
membership protocols are usually implemented by modifying the
neighboring of a given node according to a certain metric in a logical
overlay.  On the contrary, we do not aim at building any overlay, as
we only rely on the physical topology so the neighbors of a given node
are not continuously modified. Lastly, these protocols are usually
cycle-based, meaning that they impose a constant load on the network
traffic, even when the system state does not change. Likewise, this
also prevents from using solutions that periodically advertise their
cache content~\cite{garcia-lopez, hemmati2015namebased} as they also
incur an overhead even in quiet contexts without any update.
Moreover, such solutions necessarily rely on physical timeouts that
could lead to either premature removals of partitions when they are
still operating; or slow convergence where processes wrongly believe
they belong to a partition that was removed.

%Finally, we emphasize that consistency is at the core of our design,
%as \NAME provides the same guarantees for all nodes.  On the
%contrary, to avoid flooding, authors in~\cite{opnl} propose to
%propagate information about cache updates only towards the original
%replica. This enables for a request to be caught \textit{on-path},
%and be redirected to the closest replica. However, the announcement
%is only performed towards this original replica, that is requests can
%miss closer replicas that are not on this path, thus decreasing the
%efficiency of the cache sharing mechanism.

%AVK: This should go where we explain the DEL, as this is really
%specific...  One could also solve this issue by removing partitions
%that were not advertised for a defined
%time~\cite{hemmati2015namebased, garcia-lopez}.  However, relying on
%physical timeout could lead to either premature removals of
%partitions when they are still operating; or slow convergence where
%processes believe they belong to a partition that was removed
%(\TODO{plot ?}). In addition, since timeouts imply a continuous
%maintenance of partitions, such partitioning protocol incurs an
%overhead even in quiet contexts without dynamic partitions.

% AVK t-man: T-Man: Gossip-Based Overlay Topology Management5doat a
% random time at a random time once in eachconsecutive interval of T
% time unit This buffer contains a random sample of the nodes from the
% entirenetwork. It is provided by apeer sampling service We do not
% try to to create a topology aware overlay ! Because locality.  We do
% not provide routing ! routing is two determine path between any pair
% of nodes. We only provide path to objects Koala locality aware
% routing. Long links. Lazzy (nothing happens) Koala shares many core
% aspectswith similar overlays, such as Chord and Symphony.  VoroNet:
% overlay and routing. i.e. long links It links application objects
% rathen than physical nodes so that objects with similar
% characteristic are neighbors.  are logigal overlay, that are not
% correlated to the physical infrstructure.  Usually rely on long
% links...  AVK

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "../paper"
%%% ispell-local-dictionary: "english"
%%% End: