### 2021-06-09 rebuttal.tex
add rebuttal 495

We thank the reviewers for their thorough reading  and interesting comments.
We understand that althought the paper is relevant, some parts should be clarified. We would be happy to be shepperded if reviewers believe it is relevant.

Below, we reply to each remark in more detail.

- R1.1. "forwarding the best known partition"  consists in transmitting the best location where to get the replica.
It could be more formally defined.

- R1.2.  We agree ``Theorem'' is more appropriate than ``Corollary''. Regarding the proof, additional information would be given (under shepperding).

- R1.3. Regarding Theorem 2, we chose to list all cases to help understand the algorithm.
Every Add/Del operation creates a small subset of events that depends on its nature,
and the neighbors' history. This allows us to perform a neighbor-to-neighbor analysis of cases and avoid any combinatorial explosion.

- R2.1. Our objective is to introduce AS-cast and its properties in a generic manner. 
Hence, experiments focus on evaluating the correctness/costs of our protocol.
As stated in Conclusion, we plan to assess the use of AS-cast in IPFS distributed storage. This study would clarify its tradeoff depending on object size, update frequency, etc.

- R2.2. AS-cast is more efficient than CDN: it allows (i) retrieving the closest location eventually and (ii) partitioning the infrastructure dynamically according to the creation/removal of replicas. Contrarily, CDN relies on a static partitioning of the infrastructure: nodes get replicas from a subset of specific nodes and for all objects. A node may contact a peer of the CDN infrastructure while a replica may exist only one hop away (replicated by one of its neighbors). 
AS-cast natively deals with such cases as any node can serve a replica.

- R2.3.  The scalability issue related to the management of thousands of 
objects should be addressed. As stated in Conclusion, we propose to leverage the
hierarchical properties of inter-autonomous systems to avoid flooding uninterested nodes. 
However, this increases the complexity to prove that our protocol still converge.

- R3.1. In addition to  R2.3, we confirm that partitioning is per-object.
Yet, it would be straightforward indeed to mutualize traffic, e.g., using dedicated propagation windows to consolidate control messages together.

- R3.2. AS-cast does not handle replication strategies. If it exists only one source and it crashes, this reference eventually disappears from the system. AS-cast reacts similarly to crashes and departures of sources (see Algorithm 3 Line 4) to update the best partition known by impacted nodes.

- R3.3. We do not  address the issue of content mutability.
From our viewpoint, this aspect is application-dependent. For instance, mutability could be provided by versioning as done in IPFS, BlobSeer, etc.

- R3.4. Regarding the experiments remarks, please, refer to R2.2.

- R3.5. Several approaches tackle the indexation problem in NDN. We were unaware of NDNS and agree to add it in Related Work.

- R3.6. Although network topologies include triangular inequalities,  they do not endanger the AS-CAST correctness (AS-CAST relies on the shortest distance between vertices).