### 2021-09-09 input/figdel.tex
address Adrien's comments

\begin{figure}
  \begin{center}
    \subfloat[Part A][\label{fig:delA}\Process~$a$ deletes its
      partition.  It notifies all \processes that belong to its
      partition.]{\input{input/figdelA.tex}}
    \hspace{5pt}
    %
    \subfloat[Part B][\label{fig:delB}$\delta$ stops as soon as they
      encounter another partition. \Process~$b$ answers with its
      partition.]{\input{input/figdelB.tex}}
  \end{center}
  \caption{\label{fig:del}Efficient removal of a partition using
    \NAMEB. $a$ eventually knows that it belongs to $P_d$.}
\end{figure}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "../paper"
%%% ispell-local-dictionary: "english"
%%% End: