### 2021-11-16 input/figundoproblem.tex

\begin{figure*}
  \newcommand{\SCALE}{0.9}

  \newcommand{\LEFT}{\triangleleft}
  \newcommand{\RIGHT}{\triangleright}

  \begin{center}
    \subfloat[Part A][\label{fig:undoproblemA}$a$ and $c$ become sources.
      $w_{ab} = 2$; $w_{bc} = 1$.]{\input{input/figundoproblemA.tex}}
    \hspace{0pt}
    \subfloat[Part B][\label{fig:undoproblemB}($c$ deletes its partition but $a$ does not.
      $b$ delivers and forwards $\alpha_a$.]
             {\input{input/figundoproblemB.tex}}
    \hspace{0pt}
    \subfloat[Part C][\label{fig:undoproblemC}$b$ delivers and forwards $\alpha_c$.]
             {\input{input/figundoproblemC.tex}}
    \hspace{0pt}
    \subfloat[Part D][\label{fig:undoproblemD}(both) $b$ delivers and forwards $\delta_c$.]
             {\input{input/figundoproblemD.tex}}
  \end{center}
  \caption{\label{fig:undoproblem}From $c$'s perspective, this
    scenario and Figure~\ref{fig:problem} are similar. Even when $a$
    did not delete its partition, $c$ must acknowledge $P_a$'s
    \emph{possible} deletion and propagate it.}
\end{figure*}