### 2021-12-14 input/proposal.tex
reworked detection and followup text

\section{Adaptive scoped broadcast}
\label{sec:adaptive}

To provide lazy and dynamic logical partitioning in dynamic networks
using scoped broadcast, all live \processes must collaborate to
disseminate messages that notify new or removed sources to all and
only interested \processes. This section reviews step-by-step the
properties that allow \processes to converge to the desired state
together. It first defines scoped broadcast, then uses it to guarantee
consistent partitioning when a \process can only become a new source
in the system. It highlights the issue when a \process can also remove
its status of source. It shows that using local knowledge and scoped
broadcast, \processes can still reach dynamic consistent partitioning
when they are able to detect possible blocking conditions in the
dissemination of required notifications. Finally, it further constrain
the system by allowing only subsets of \processes to reach lazy
consistent partitioning. This improvement only requires \processes to
know which of their neighbors belong to their own network.

\subsection{Scoped broadcast}
\label{subsec:scoped}

In this paper, we consider Edge infrastructures as a set of
interconnected autonomous systems comprising heterogeneous \nodes
interconnected by communication links. \Processes involved in the
management of content may crash but are not byzantine.  \Processes can
reliably communicate through asynchronous message passing to other
known \processes called neighbors.  We define scoped broadcast as a
communication primitive that propagates a message around its
broadcaster within an application-dependant scope.

\begin{definition}[Autonomous system]
  An autonomous system is a network comprising nodes and communication
  links that we represent as a \underline{g}raph of
  \underline{v}ertices and \underline{e}dges:
  $G = \langle V, E \rangle$ with $E \in V \times V$. A
  \underline{p}ath $\pi_{xz}$ from \Process~$x$ to \Process~$z$ is a
  sequence of contiguous edges
  $[\langle x, y_1 \rangle, \langle y_1, y_2\rangle, \ldots, \langle
  y_n, z \rangle]$.
\end{definition}

\begin{definition}[\label{def:scoped}\underline{S}coped broad\underline{cast} (\NAMEB)]
  When \Process~$x$ scoped \underline{b}roadcasts $b_x(m)$ a
  \underline{m}essage $m$, every correct \process $y$ within a scope
  \underline{r}eceives $r_y(m)$ and \underline{d}elivers it
  $d_y(m)$. The scope depends on the \underline{s}tate $\sigma$ of
  each \process, the \underline{m}etadata $\mu$ piggybacked by each
  message, and a \underline{p}redicate $\phi$ verified from \process
  to \process:
  $(b_x(m) \wedge r_y(m)) \implies \exists \pi_{xy}: \forall z \in
  \pi_{xy}, \phi(\mu_z, \sigma_z)$.
\end{definition}

This definition encompasses more specific definitions of related
work~\cite{hsiao2005scoped, lue2006scoped, wang2015prodiluvian}. It
underlines the transitive relevance of messages. It also highlights
that the functioning of epidemic propagation is well-aligned with the
objectives of scoped broadcast. As consequence, we assume
implementations relying on the forwarding of messages from
neighbor-to-neighbor.

\begin{definition}[\label{def:forwarding}Forwarding]
  A \process $x$ \underline{f}orwards a message $m$ by
  \underline{s}ending it to its neighbors $y$ with metadata that
  depends on its local knowledge: $f_x(m) \implies \forall \langle x,
  y\rangle \in E: s_{xy}(m \TODO{\oplus} \sigma_x)$.
\end{definition}

We use scoped broadcast to efficiently modify the state of each
\process depending on the partitions that exist in the system.

\subsection{Consistent partitioning}
\label{subsec:consistent}

At any time, a \process can decide to become a \emph{source}, hence
creating a new partition in the system by executing an \texttt{Add}
operation. This partition includes at least its source plus
neighboring \processes that are closer to this source than any other
one. Such a distance (or \emph{weight}) is application-dependant: in
the context of maintaining distributed indexes, this would be about
link latency that \nodes could monitor by aggregating \texttt{ping}s;
or operational costs when dealing with multiple tenants.

\input{input/figadd.tex}

%% \begin{definition}[\label{def:partitioning}Partitioning]
%%   Let $S \subseteq V$ be the set of \underline{s}ources, and $P_{s\in
%%     S}$ be the \underline{p}artition including at least \Process~$s$,
%%   each \process belongs to at most one partition $\forall x, y \in V,
%%   \forall s,s' \in S: (x \in P_{s} \wedge y \in P_{s'}) \implies (x \neq
%%   y \vee s = s')$, and there exists at least one path $\pi_{xs}$ of
%%   \processes that belong to this partition $\forall z \in \pi_{xs}: z
%%   \in P_s$.
%% \end{definition}

%% Definition~\ref{def:scoped} and Definition~\ref{def:partitioning}
%% share the transitive relevance of \process states. However, we further
%% constrain the partitioning in order to guarantee the existence of
%% exactly one consistent partitioning that \processes eventually
%% converge to.

\begin{definition}[\underline{C}onsistent \underline{p}artitioning (CP)]
  Assuming a set of \underline{s}ources $S\subseteq V$, a positive
  \underline{w}eight $w_{xy}$ associated with each edge $\langle x, y
  \rangle \in E$, we define consistent partitioning as a set of
  logical partitions $P_{s\in S}$ where each \node $x$ belongs to the
  partition of its closest source $s$, \ie there exists a
  \underline{p}ath $\pi_{sx}$ with a sum of weights $|\pi_{sx}| =
  \sum\{w_{pq} | \langle p, q \rangle \in \pi_{sx}\}$ smaller than any
  other path, with $|\pi_{xx}|$ being $x$'s greatest lower bound.
  %$\forall x \in
  % P_{s}:  \nexists P_{s'}$ such that $|\pi_{s'x}| < |\pi_{sx}|$.
\end{definition}

Unfortunately, \processes do not share a common global knowledge of
the network state. For \processes to eventually ($\eventually$) reach
consistent partitioning, each source $s$ must send
\texttt{\underline{A}dd} notifications $\alpha_s$ to all \processes
that are closer to it than any other source, along with enough
metadata to allow them to decide of their closest source. A simple
implementation based on Definition~\ref{def:scoped} consists in
ensuring that each \process $x$ eventually receives
$\alpha_s^{|\Pi_{sx}|}$ where $\Pi_{sx}$ is the \emph{best} path from
any source to it; the best value $|\Pi_{sx}|$ being built over
forwarding along this path.

\begin{theorem}[\label{theo:fb}\underline{F}orwarding of \underline{B}est (FB)
    $\implies\eventually \textup{CP}$]
    % 
    Assuming that each \process $x$ stores its outgoing weights
    ($\forall \langle x, y \rangle \in E: \pi_{xy} \in \sigma_x$), a
    total order on messages based on weights ($m^d_s < m^{d'}_{s'}$
    when $d < d' \vee (d = d' \wedge s< s')$), reliable communication
    links ($s_{xy}(m) \iff \eventually r_{yx}(m)$), \processes eventually reach
    consistent partitioning if each \process delivers its best (\ie
    smallest) message among \underline{r}eceived messages $R_x$, and
    forwards its best message among \underline{d}elivered messages
    $D_x$ such that $f_x(\alpha^d_s) \implies \forall \langle x, y
    \rangle \in E: s_{xy}(\alpha_s^d \oplus ^{w_{xy}})$, \ie by
    accumulating the respective weight of used edges.
\end{theorem}

\begin{proof}
  Whenever a \process~$s$ becomes a source, it broadcasts hence
  delivers its own message $d_s(\alpha_{s}^{|\pi_{ss}|})$. Whatever
  its set of received messages, it acknowledges that it belongs to its
  own partition $P_s$ since $\alpha_{s}^{|\pi_{ss}|} = \min D_s$ and
  it remains forever since $|\pi_{ss}|$ is its greatest lower bound.
  \noindent Such a source forwards its notification to its neighbors.
  Every neighbor eventually receives its notification since
  communication links are reliable. Whatever the order of received
  messages $R_x$ at neighboring \process~$x$, total order ensures that
  it delivers notifications $\alpha_s^{d}$ when $d= |\pi_{ss}| +
  w_{sx} = \min |\pi_{s'x}|$ where $\min |\pi_{s'x}|$ is the lightest
  weight of $x$ to any source $s'$.
  \noindent Among these neighbors, at least those that fulfill the
  latest condition forward their respective notification.  By
  transitivity, the message originating from $s$ reaches all
  \processes that belong to $P_s$ at least through their respective
  lightest path: $\small\smash{\forall y \in V, s, s' \in S: \min
    |\pi_{s y}| < \min |\pi_{s' y}| \implies \eventually
    d_{y}(\alpha_{s}^{\min |\pi_{sy}|})}$. When the system becomes
  quiescent, \ie no \process becomes source anymore, every \process
  eventually acknowledges the partition it belongs to, \ie \processes
  eventually reach consistent partitioning together. In addition, the
  protocol terminates: a \process never delivers hence forwards a
  message after it received, delivered, and forwarded the message of
  its closest source from its lightest path.
\end{proof}

\begin{algorithm}
  \input{input/algoadd.tex}
  \caption{\label{algo:add}Add-only CP protocol at \Process~$p$.}
\end{algorithm}

Algorithm~\ref{algo:add} shows the few instructions that implement a
\emph{reactive} forwarding of best to reach consistent partitioning in
a static network where \processes never join, crash, nor leave the
system. As soon as a \process receives a better message, it delivers
and forwards it. The trade-off consists in decreasing termination time
at cost of increased traffic.  Figure~\ref{fig:add} illustrates the
behavior of this algorithm on a system comprising 4 \processes $a$,
$b$, $c$, and $d$. Both $a$ and $d$ become sources.  They \NAMEB
notifications $\alpha_a^0$ and $\alpha_d^0$. They initialize their own
state with the lowest value $0$ (see Line~\ref{line:lowestbound}), and
send a message to each of their neighbors by accumulating the
corresponding edge weight (see Line~\ref{line:accumulator}). In
Figure~\ref{fig:addC}, $b$ receives $\alpha_{d}^{1}$. Since it
improves its own partition distance, it keeps it and forwards it to
its neighbors. In Figure~\ref{fig:addD}, $b$ discards
$\alpha_{a}^{2}$, for it does not improve its partition
distance. Neither $c$ nor $d$ will ever acknowledge that Source~$a$
exists.  The protocol lacks of obvious traffic optimization, \eg grey
messages are useless in this scenario. Nevertheless, the system
discards last transiting messages after it reached consistent
partitioning.

Each \process has a space complexity of $\mathcal{O}(\log_2(d \cdot
|V|))$, and generates messages of identical space complexity. In terms
of number messages, in the average case, a \process $x$ chosen
uniformly at random among all \processes creates a logical
partition. Its messages $\alpha_x$ spread through the network until
reaching \processes that belong to another partition closer to
them. This splits partitions in half on average. Overall, the $n^{th}$
new partition comprises \smash{$\mathcal{O}(\frac{|V|}{2^{\lfloor
      \log_2 n \rfloor}})$} \processes. This decreases every new
partition until reaching one \process per new partition: its
source. The average number of messages per \process is
\smash{$\mathcal{O}(\frac{\overline{|O|}}{2^{\lfloor \log_2 n
      \rfloor}})$}, where \smash{$\overline{|O|}$} is the average
number of neighbors per \process.

While only adding logical partitions to the distributed system is
straightforward and lightweight, removing them introduces additional
complexity and costs caused by concurrent operations.

\subsection{Dynamic consistent partitioning}
\label{subsec:dynamic}

At any time, a source can revoke its status of source by executing a
\texttt{Del} operation, hence deleting its partition from the
system. All \processes that belong to this partition must eventually
choose another partition to belong to.

\input{input/figsimplerproblem.tex} %% positioning

\input{input/figdel.tex}

\begin{definition}[\label{def:dcp}\underline{D}ynamic
    \underline{c}onsistent \underline{p}artitioning (DCP)] A DCP
  protocol enables \processes to join or leave the set of sources at
  any time while ensuring eventual consistent partitioning.
  % Dynamic consistent partitioning is consistent partitioning where
  % \processes can join or leave the set of sources at any \TODO{time}.
\end{definition}

Delete operations bring a new notion of order between events, and most
importantly between message deliveries. A \process $s$ that performs a
\texttt{Del} operation implicitly states that all its preceding
\texttt{Add} operations become obsolete. As consequence, their results
-- scoped broadcast in the form of \texttt{\underline{A}dd}
notifications $\alpha_s$ -- should be canceled by the corresponding
\texttt{\underline{D}el} notifications $\delta_s$. We focus on the
last delivery of each \process, since the best is also the last, as
highlighted by Algorithm~\ref{algo:add}.

\begin{definition}[Happens-before ($\rightarrow$)~\cite{lamport1978time}]
  The transitive, irreflexive, and antisymmetric happens-before
  relationship defines a strict partial order between events. Two
  messages are concurrent if none happens before the other.
\end{definition}

\begin{definition}[\label{def:lwo}Stale messages]
  Only the latest broadcast of a \node matters.  A message $m$ conveys
  \underline{s}tale control information $\mathcal{S}(m)$ as soon as
  its broadcaster broadcasts another message: $\mathcal{S}(\alpha_x) =
  \exists \delta_x: b_x(\alpha_x) \rightarrow b_x(\delta_x)$.  A
  \process only delivers or sends messages that it assumes up-to-date.
  % : $s_x(\alpha_s)
  % \implies \delta_s \not\in R_x$.
  For convenience, we define
  % \TODO{$\textup{minUTD(}D\textup{)} = \min
  %  \{\alpha_s^d | \alpha_s^d \in D \wedge \delta_s \notin D\}$}, and
  $\textup{last(}D\textup{)} = \alpha_s \in D: \nexists \alpha_{s'}
  \in D: d_x(\alpha_s) \rightarrow d_x(\alpha_{s'})$.
  %  \TODO{Only keep last?  Tothink what happens if we keep all as in
%    def now.}
\end{definition}

\newcommand{\last}{\textup{last}~}

A naive attempt at implementing DCP resembles echoes in
acoustics. Following the principles of scoped broadcast as stated in
Definition~\ref{def:scoped}, a \process that receives a staleness
notification forwards it only if the latter targets its current
partition. The staleness notifications propagate through the network
as long as they remain in the deleted partition. When they reach the
bordering \processes of the deleted partition, it creates an echo of
bordering partitions that will go backward to fill the gap left open
by removals using forwarding of best (FB) as stated by
Theorem~\ref{theo:fb}.

\input{input/figundoproblem.tex}

\begin{definition}[\label{def:fs}\underline{F}orwarding of
  \underline{s}taleness (FS)]
  %
  Any source can broadcast a staleness notification at any
  time. Every \process $x$ delivers and forwards a received staleness
  notification if it targets its best up-to-date delivered message:
  %% \TODO{$(\delta_s \in R_x \wedge \alpha_s =
  %%     \textup{minUTD}~D_x) \iff \eventually (d_x(\delta_s) \rightarrow
  %% f_x(\delta_s))$.}
  $(\last D_x = \alpha_s \wedge d_x(\alpha_s) \rightarrow
  r_x(\delta_s)) \implies d_x(\delta_s) \rightarrow f_x(\delta_s)$.
  % \wedge \nexists d'_s: d_x(\delta_s) \TODO{\not\rightarrow} d_x(\delta_s'))$.
\end{definition}

\begin{definition}[\label{def:fbplus}$\textup{FB}^+$: echos] FB where
  a \process $x$ that receives but does not deliver a staleness
  notification sends back --~or echoes~-- its best up-to-date
  delivered message: $(\last D_x = \alpha_s^d \wedge \delta_s
  \not\in D_x \wedge d_x(\alpha_s^d)\rightarrow r_{xy}(\delta_{s'}))
  \implies r_{xy}(\delta_{s'}) \rightarrow
  s_{xy}(\alpha_s^d\oplus^{w_{xy}})$.

  % $r_{xy}(\alpha_s)
  % \wedge \delta_s \in D_x \implies s_{xy}(\alpha_{s'}^d\oplus
  % ^{w_{xy}})$ with $\alpha_{s'^d} = \min D_x \wedge \delta_{s'} \not\in
  % D_x$. \TODO{Rework min}
\end{definition}

Figure~\ref{fig:del} illustrates the behavior of this implementation
($\textup{FB}^+ \wedge \textup{FS}$).  Two partitions initially exist:
$P_a$ and $P_d$ that respectively include $\{a\}$, and $\{b, c,
d\}$. In Figure~\ref{fig:delA}, $a$ deletes its partition. FS in
Definition~\ref{def:fs} states that $a$ must notify all its
neighbors~--~here only $b$~-- that may belong to its partition.  In
Figure~\ref{fig:delB}, $b$ receives but does not deliver $\delta_a$
since $\delta_a$ does not target its current partition $P_d$.
$\textup{FB}^+$ in Definition~\ref{def:fbplus} states that $b$ must
echo back its own best up-to-date message $\alpha_d^3$, for it may be
the best for $a$. Eventually, every \process belongs to Partition
$P_d$. In this scenario, they reach consistent
partitioning. Unfortunately, this protocol does not ensure consistent
partitioning in every scenario.

% \TODO{The naive implementation described above is $\textup{FB}^+
%  \wedge FS$. And it does  not work.}

\begin{lemma}[\label{lem:fbfse}$\textup{FB}^+ \wedge \textup{FS}
    \centernot\implies \textup{DCP}$] Forwarding best up-to-date,
  forwarding staleness, and echoing is not sufficient to guarantee
  dynamic consistent partitioning.
\end{lemma}

\begin{proof}
Stale control information as stated in Definition~\ref{def:lwo} may
impair the propagation of both
\begin{inparaenum}[(i)]
\item notifications about actual sources, and
\item notifications about deleted partitions.
\end{inparaenum}
Figure~\ref{fig:problem} depicts a scenario comprising only three
\processes $a$, $b$, $c$ chained with FIFO links, \ie where \processes
receive the messages in the order of their sending ($s_{pq}(m)
\rightarrow s_{pq}(m') \implies r_q(m) \rightarrow r_q(m')$). In
Figure~\ref{fig:problemA}, both $a$ and $c$ become sources, sending
their respective notification message to $b$. In
Figure~\ref{fig:problemB}, both $a$ and $c$ delete their partition
while $b$ receives, delivers, and forwards $\alpha_a^2$. In
Figure~\ref{fig:problemC}, $b$ receives, delivers, and forwards
$\alpha_c^1$, for it improves its best partition. Then it receives and
discards $\delta_a$, for its best partition does not match the deleted
one. We simplify the echo since $\alpha_c^3$ is already transiting
towards $a$. Finally, Figure~\ref{fig:problemD} shows that eventually,
$c$ stays in the deleted partition $P_a$ for not having received
$\delta_a$ that $b$ blocked earlier. The system does not reach
consistent partitioning. \end{proof}

As a cascading effect in Figure~\ref{fig:problemD}, not only $c$
wrongfully believes that it belongs to $a$'s partition when $a$
already deleted it, but it could contribute to its inconsistency by
blocking farther but up-to-date messages from actual sources.

A first step to avoid staying in inconsistent state is to extend the
behavior of each \process so \processes such as $c$, that would remain
in a wrong partition, use their history of receipt and delivery to
detect the blocking of staleness notifications that can hinder
reaching consistent partitioning. For instance, $c$ detects the
blocking of $\delta_a$ because it acknowledges that its parent $b$
delivered a stale notification $\alpha_c^1$, \ie between the delivery
of $\alpha_c^1$ and the delivery of $\delta_c$, $b$ blocked any other
$\delta$.

%% \Begincomme{inparaenum}[(i)] \item to detect possibly blocking
%% conditions that can hinder \item to purge all stale messages from
%% the system \item enabling up-to-date notifications to eventually
%% echo back to every \process.  \end{inparaenum} Since
%% Figure~\ref{fig:problemD} highlights that when the protocol
%% terminates, $c$ is the \process that remains in the wrong
%% partition, we put the burden of detection and resolution on it.

%% \TODO{A lot of blockings occur, only that of example is a problem.
%%   Could be $b$ or $c$ that detects? We put burden on $c$ since
%%   unconsistency starts here. Problem: how to make sure that $c$
%%   receives a staleness notification in this case? Also, $b$ is not
%%   sure to receive $\delta_c$ since composable problem.}

\begin{lemma}[\label{lem:detector}$\textup{FB}^+ \wedge \textup{FS}
  \implies$ Detector existence] Assuming
  $\textup{FB}^+$ and $\textup{FS}$, a chain of delivery on $\pi_{xz}$ of
  $\alpha_x$ with $\alpha_x = \last D_z$, if $x$ broadcasts a staleness
  notification $\delta_x$, either $z$ eventually delivers it, or a
  \process $y$ in such a chain of delivery $\pi_{xz}'$ eventually
  acknowledges its blocking.
  % $b_p(\delta_x) \implies d_q(x) U Detect[X]$ ??? 
\end{lemma}

\begin{proof}
  %% knowing that b_p delta is sufficient to say that processes will
  %% receive delta or detect blocking
  (A) Assuming a chain of delivery $\pi_{xz} = [x, \ldots, z]$ of
  $\alpha_x$ with $\forall y \in \pi_{xz}: \alpha_x = \last D_y$, and
  assuming $d_x(\delta_x)$, then with forwarding of staleness
  (Definition~\ref{def:fs}),
  $f_{\pi_{xz}[k]}(\delta_x) \implies \eventually
  d_{\pi_{xz}[k+1]}(\delta_x)$ except if $\delta_x \in D_{\pi_{xz}[k+1]}$ which
  means that $\pi_{xz}[k+1]$ already delivered and forwarded $\delta_x$
  following another such a chain of delivery. Therefore, $z$
  eventually receives, hence delivers $\delta_x$.

  \noindent (B) Assuming a chain of delivery
  $\pi_{xz} = [x, \ldots, z]$ of $\alpha_x$ with at least
  $\alpha_x = \last D_z$, and assuming $d_x(\delta_x)$, then by (A),
  the staleness notification reaches a \process $ y = \pi_{xz}[k]$. By
  Definition~\ref{def:fs}, the forwarding stops when $y$ already
  delivered $\delta_x$ ($\delta_x \in D_y$ covered by (A)) or
  delivered a better message $\alpha_s = \last D_y$ with
  $\alpha_s \neq \alpha_x$. With $y'=\pi_{xz}[k+1]$, forwarding of
  best (Definition~\ref{def:fbplus}) implies three possible outcomes:
  \begin{enumerate}[(i)]
  \item $\alpha_{s'} = \last D_{y'}$ with $\alpha_{s'} \neq \alpha_x$
    hence $y' \iff y$ which leaves two possible outcomes:
  \item
    $r_{y'}(\alpha_s) \wedge \alpha_x = \last D_{y'} \wedge \alpha_x <
    \alpha_s$ which means that a shorter path of delivery $\pi_{xz}'$
    exists that either forwards appropriate staleness notifications
    (A) or detects its possible blocking:
  \item
    $r_{y'}(\alpha_s) \wedge \delta_s \in D_{y'} \wedge \alpha_x =
    \last D_{y'}$ which detects a possible blocking of
    $\delta_x$. Without global knowledge, $y'$ assumes it belongs to
    the shortest and only path of delivery of $\alpha_x$
    ($\pi_{xy'} \subseteq \Pi_{xz}$) thus it does not further delegate
    the detection to another \process.
  \end{enumerate} \vspace{-1.5em} %% ugly vspace but necessary for enumerate+qed
  %
  %% (i) If every \process (such as $a$) starting from the source
  %% delivers and forwards a removal notification $\delta_X$ when their
  %% last delivery is $\alpha_X$, then every such \process eventually
  %% delivers the removal notification $\delta_X$ except \processes (such
  %% as $b$) that delivered $\alpha_X$ from a \process that delivered a
  %% message from another partition $\alpha_Y$ since then.
  %
  %% These exceptions eventually receive and deliver $\alpha_Y$ since
  %% $\alpha_Y$ is better than $\alpha_X$ through this path, except if
  %% they already received and delivered the removal notification of
  %% $P_Y$ (such as $c$). These \processes may never receive hence
  %% deliver $\delta_X$, and may never receive hence deliver a better
  %% message than $\alpha_X$. They need an additional mechanism to
  %% eventually purge $\alpha_X$ that cannot rely on the eventual purging
  %% of $P_Y$ at \processes like $b$, to avoid deadlocks.
  %
  %% (ii) Assuming that every \process keeps an history of its past
  %% deliveries, \processes (such as $c$) can detect the inconsistency,
  %% since they receive from their parent an already deleted partition
  %% $P_Y$. This notification means that the parent discards any
  %% $\delta_Z$ with $\alpha_Z^z < \alpha_Y^y$, and most importantly, if
  %% $\delta_X$ exists, it discards it. To ensure the purge of stale
  %% messages, a detecting \process must assume the worst case that such
  %% $\delta_X$ exists, and send another kind of message, noted $\Delta$,
  %% that notify the possible removal of $P_X$.
  %
  %% (iii) A \process (such as $d$) whose last delivery is $\alpha_X$,
  %% but whose parent is neither inconsistent (like $b$) nor receiving
  %% $\delta_X$ (like $a$), eventually receives $\Delta$ from a detecting
  %% \process, either directly or transitively, for such a child \process
  %% ($d$) trusts the possible removal of $P_X$ by delivering and
  %% forwarding such $\Delta$. $\Delta$ suffers from identical blocking
  %% conditions (between $b$ and $c$) than $\delta$, leading to the same
  %% solutions of detection and propagation of $\Delta$. Eventually,
  %% every \process whose last delivery is $\alpha_X$ receives and
  %% delivers either a better $\alpha_Z$, or $\delta_X$, or $\Delta_X$.
\end{proof}

As a second step, detecting \processes must proactively purge the
system from their forwarded notifications. For instance in
Figure~\ref{fig:problemD}, $c$ could broadcast $\delta_a$ in order to
acknowledge the staleness of its own $\alpha_a^3$ and inform
neighboring \processes that delivered it as well. Unfortunately, proof
of Lemma~\ref{lem:detector} shows that there exists false positive in
the blocking detection: $\delta_x$ may not exist but $y'$ still
receives $\alpha_s$ after the delivery of $\delta_s$ from the path it
delivered $\alpha_x$. Figure~\ref{fig:undoproblem} highlights this
behavior. In Figure~\ref{fig:undoproblemB}, $a$ does not delete its
partition. In Figure~\ref{fig:undoproblemD} and because of scoped
broadcast, $c$ received the same series of messages as in
Figure~\ref{fig:problemD}. Yet, $c$ must assume the existence of
$\delta_a$ and act accordingly by forcing the staleness of its best
delivered message and disseminate this information to its neighbors.
To avoid flooding the system with false positive staleness
notifications, we propose to reduce the scope of staleness
notifications by including only downstream \processes. In
Figure~\ref{fig:undoproblemD}, false positives would not affect $a$
and $b$.

\begin{definition}[\label{def:fsplus}$\textup{FS}^+$: forwarding
    staleness downstream] Any \emph{\process} can broadcast a
  staleness notification at any time. A \emph{child \process} $x$
  delivers and forwards a received staleness notification if it comes
  from the path of delivery of its best up-to-date delivered
  message.
\end{definition}

It is worth noting that forwarding of staleness as stated in
Definition~\ref{def:fs} becomes a specific case of
Definition~\ref{def:fsplus} where the source itself forwards a
staleness notification downstream. The notification must reach all
\processes in its partition since the source has no \processes
upstream, and belongs to the delivery path of all \processes that
delivered this message.

\begin{theorem}[$\textup{FB}^+ \wedge \textup{FS}^+ \implies \textup{DCP}$]
  A protocol guarantees dynamic consistent partitioning if it
  implements forwarding of best up-to-date messages with echos,
  forwarding of staleness messages downstream, and the detection of
  possible blocking triggers the forwarding of staleness notifications
  downstream.
\end{theorem}

\begin{proof}
  Detection triggers forwarding of staleness downstream which
  completes the case study of Lemma~\ref{lem:detector} by ensuring
  that, when a source broadcasts its staleness notification, all
  \processes that belong to its partition eventually deliver such a
  staleness notification. All downstream bordering \processes also
  eventually receive such a staleness notification and echo back their
  best delivered message. This triggers another competition as in
  Theorem~\ref{theo:fb} for the \processes that delivered the
  staleness notification.
\end{proof}

\begin{algorithm}
  \input{input/algoascast.tex}
  \caption{\label{algo:ascast}\NAME: DCP protocol at \Process~$p$.}
\end{algorithm}

Algorithm~\ref{algo:ascast} shows the instructions of our
implementation called \NAME that enables dynamic consistent
partitioning. It implements forwarding of best (see
Line~\ref{line:fb}) and echos (see Line~\ref{line:echo}). To implement
forwarding of staleness downstream as stated in
Definition~\ref{def:fsplus},
\begin{inparaenum}[(A)]
\item each \process maintains a vector of versions that associates the
  respective known local counter of each known source, or has-been
  source. It constitutes a summary of known progress of other
  \processes; 
\item each notification message $\alpha$ carries the list of
  identifier and counter of each \node that forwarded it. In the worst
  case, both these structures grow linearly with the number of
  \processes in the system $\mathcal{O}(V)$. Nevertheless, following
  the principles of scoped broadcast, we expect that \processes only
  acknowledge a small subset of sources and messages;
\item each staleness notifications $\delta$ only carry the identifier
  and counter of the \process~--~source or detector~--~that generated
  it. Only downstream \processes may deliver such message, since they
  carry the identifier of the generator. 
\end{inparaenum}
To implement the detection as stated in Lemma~\ref{lem:detector}, each
\process only requires to know the direct parent of its best delivered
message which is already included in the piggybacked path of this
message. Receiving a message known to be stale from this parent
triggers the generation of staleness notifications that force
downstream \processes to reevaluate the partition they are in (see
Line~\ref{line:detect}).

\begin{algorithm}[h]
  \input{input/algoedges.tex}
  \caption{\label{algo:edges}\NAME at \Process~$p$ in dynamic networks.}
\end{algorithm}

By reusing \NAME's default behavior, Algorithm~\ref{algo:edges}
enables dynamic consistent partitioning in dynamic networks where
\processes can join, leave, or crash at any time.  Adding a \process
is equivalent to add as many edges as necessary. Since adding an edge
may improve shortest paths, the protocol triggers a competition using
echos of Definition~\ref{def:fbplus}.  Removing a \process is
equivalent to removing all its edges. Removing an edge between two
\processes may invalidate the shortest path of one of involved
\processes plus downstream \processes, or impair the propagation of
staleness notifications. Therefore, the protocol reuses its detection
mechanism of Lemma~\ref{lem:detector} as a solution.
Algorithm~\ref{algo:edges} even extends \NAME to dynamic networks
subject to physical partitioning.

% \TODO{Multiple receipt and multiple delivery}
In terms of number of messages, deleting the $a^{th}$ partition
generates twice as many messages as creating the $a^{th}$ partition:
$\delta$ messages travel through the partition, then $\alpha$ messages
compete to fill the gap \TODO{which is?}. \TODO{What about false
  positives?} Overall, the communication complexity shows that \NAME
scales well to the number of partitions.
%% In the
%% worst-case, every new partition includes all but one \process
%% belonging to the previous partition. The total number of messages
%% after the $a^{th}$ new partition is $\mathcal{O}(\overline{|O|}\cdot
%% a^2)$. As for the average-case, the number of messages for the
%% partition deletion is identical to the number of messages of the
%% corresponding partition creation.

\subsection{Lazy dynamic consistent partitioning}
\label{subsec:lazy}

\TODO{In Edge infrastructures comprising tens of thousands of autonomous
systems~\cite{nur2018crossas}, only a small subset of autonomous
systems needs logical partitioning. At any time, an autonomous system
start logical partitioning when a \process from this network becomes a
source.  Then every \process from this network belongs to the
partition of its closest source not only from their own network but
that of adjacent networks that started logical partitioning as well.
At any time, an autonomous system stops logical partitioning when the
last source from this network revokes its status of
source. Eventually, all nodes from this network do not belong to any
partition and the generated traffic stops.}

\TODO{Further scale, follow up of motivations.}

\begin{definition}[Edge infrastructure]
  An Edge infrastructure $\mathcal{G}$ has multiple autonomous systems
  interconnected with additional links $\mathcal{E} \in V_1 \times
  V_2$ where $V_1 \neq V_2$, hence $\mathcal{G} = \langle \{G_1,
  \ldots G_n\}, \mathcal{E} \rangle$. A \emph{gateway} $g$ is a
  \process with at least one link towards another network: $\langle g,
  \_ \rangle \in \mathcal{E}$.
\end{definition}

\begin{definition}[Lazy dynamic consistent partitioning (LDCP)]
  \TODO{If there exists at least a source in a network, its nodes must
    belong to the partition of the closest partition in adjacent
    networks. Otherwise, nodes must eventually belong to no
    partition.}
\end{definition}

To enable LDCP in an Edge infrastructure, we use the property provided
by DCP in an autonomous system to all its \processes and among others,
to its gateways.

\begin{lemma}[Source existence]
  When a system becomes quiescent, DCP guarantees that every \process
  eventually acknowledges the existence or non-existence of a source
  in its system.
\end{lemma}

\begin{proof}
  By design.
\end{proof}

The gateways become responsible for the flooding or the stopping of
notifications foreign to their respective network. They must
differentiate between local and global notifications. Local
notifications provide both \begin{inparaenum}[(i)]
\item local source existence, and
\item closest local source in the network
\end{inparaenum}; while global notifications only improve on closest global sources,
\ie it does not need to propagate when the foreign source is farther
than the local one.

\begin{definition}[\underline{B}roadcasting of \underline{u}ninterested (BU)]
  \TODO{meow}
\end{definition}

\TODO{Blocking condition local global remains.}

\TODO{Added invariant: global is always smaller than local when they
  exist.}

\begin{algorithm}[h]
  \input{input/algoxascast.tex}
  \caption{\label{algo:xascast}\NAMEC at \Process~$p$. \TODO{TODO.}}
\end{algorithm}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "../paper"
%%% ispell-local-dictionary: "english"
%%% End: