### 2021-12-21 input/figsimplerproblem.tex
all figures completed with what happens

\begin{figure*}
  \newcommand{\SCALE}{0.9}

  \newcommand{\SMSG}{\tiny}
  \newcommand{\OACK}{0.5}

  \thickmuskip=0mu
  \medmuskip=0mu
  \thinmuskip=0mu

  \newcommand\X{50pt}
  \newcommand\Y{-50pt}

  \newcommand{\LEFT}{\triangleleft}
  \newcommand{\RIGHT}{\triangleright}

  \begin{center}
    \subfloat[Part A][\label{fig:problemA}Both $a$ and $c$ become sources.
      $w_{ab} = 2$; $w_{bc} = 1$.]{\input{input/figsimplerproblemA.tex}}
    \hspace{5pt}
    \subfloat[Part B][\label{fig:problemB}Both $a$ and $c$ delete their partition 
     while $b$ delivers and forwards $\alpha_a$.]
             {\input{input/figsimplerproblemB.tex}}
    \hspace{5pt}
    \subfloat[Part C][\label{fig:problemC}$b$ blocks the only transiting
      $\delta_a$ while $b$ delivers and forwards $\alpha_c$.]
             {\input{input/figsimplerproblemC.tex}}
    \hspace{5pt}
    \subfloat[Part D][\label{fig:problemD}\TODO{$b$ blocks the only transiting
      $\delta_a$ while $b$ delivers and forwards $\alpha_c$.}]
             {\input{input/figsimplerproblemD.tex}}
    \hspace{5pt}
    \subfloat[Part E][\label{fig:problemE}\TODO{$b$ blocks the only transiting
      $\delta_a$ while $b$ delivers and forwards $\alpha_c$.}]
             {\input{input/figsimplerproblemE.tex}}
    \hspace{5pt}
    \subfloat[Part F][\label{fig:problemF}$b$ delivers and forwards $\delta_c$. $c$
    stays in the deleted partition $P_a$.]
             {\input{input/figsimplerproblemF.tex}}
  \end{center}
  \caption{\label{fig:problem}Even in the simplest scenarios, naive
    propagation of $\alpha$ and $\delta$ messages may be insufficient
    to guarantee consistent partitioning. In the bottom scenario, if
    $c$ had children, they would stay in the wrong partition
    too.
    %
    % Despite both scenarios look the same from $c$'s perspective, $c$
    % must acknowledge $P_a$'s \emph{possible} deletion and propagate
    % it.
  }
\end{figure*}