### 2021-12-21 input/figundoproblem.tex

\begin{figure*}
  \newcommand{\SCALE}{0.9}

  \thickmuskip=0mu
  \medmuskip=0mu
  \thinmuskip=0mu

  \newcommand\X{50pt}
  \newcommand\Y{-50pt}

  \newcommand{\OACK}{0.5}
  \newcommand{\SMSG}{\tiny}

  \newcommand{\LEFT}{\triangleleft}
  \newcommand{\RIGHT}{\triangleright}

  \begin{center}
    \subfloat[Part A][\label{fig:undoproblemA}$a$ and $c$ become sources.
      $w_{ab} = 2$; $w_{bc} = 1$.]{\input{input/figundoproblemA.tex}}
    \hspace{5pt}
    \subfloat[Part B][\label{fig:undoproblemB}$c$ deletes its partition but $a$ does not.
      $b$ delivers and forwards $\alpha_a$.]
             {\input{input/figundoproblemB.tex}}
    \hspace{5pt}
    \subfloat[Part C][\label{fig:undoproblemC}$b$ delivers and forwards $\alpha_c$.]
             {\input{input/figundoproblemC.tex}}
    \hspace{5pt}
    \subfloat[Part D][\label{fig:undoproblemD}$b$ delivers and forwards $\delta_c$.]
             {\input{input/figundoproblemD.tex}}
    \hspace{5pt}
    \subfloat[Part E][\label{fig:undoproblemE}$a$ echoes back $\alpha_a^2$ that $b$ needs.]
             {\input{input/figundoproblemE.tex}}
    \hspace{5pt}
    \subfloat[Part F][\label{fig:undoproblemF}$a$, $b$, and $c$ belong to $P_a$.]
             {\input{input/figundoproblemF.tex}}
  \end{center}
  \caption{\label{fig:undoproblem}From $c$'s perspective,
    Figure~\ref{fig:problemE} and Figure~\ref{fig:undoproblemE} are
    similar in terms of received messages, but the outcomes eventually
    differ. Yet, $c$ must act on Figure~\ref{fig:undoproblemE}, and
    acknowledge then propagate the \emph{possible} staleness of its
    partition $P_a$.}
\end{figure*}