### 2022-01-28 input/conclusion.tex
final draft for conclusion

\section{Conclusion}
\label{sec:conclusion}

With the advent of the Edge and IoT era, major distributed services
proposed by our community should be revised to mitigate as much as
possible traffic between \processes.  In this paper, we addressed the
challenge of content indexing as a dynamic logical partitioning
problem where partitions grow and shrink to reflect content and
replicas locations, as well as infrastructure changes.  Using an
efficient communication primitive called scoped broadcast, each
\process composing the infrastructure eventually knows the \process
hosting the best replica of a specific content.  The challenge resides
in handling concurrent operations that may impair the propagation of
messages, and in turn, lead to inconsistent partitioning.
%% The major difficulty lies in concurrent operations, each having a
%% different scope, and ensure the correct convergence of the system.  We
%% defined, in particular, multiple properties that highlight this
%% complexity.
%
We highlighted the properties that solve this problem and proposed an
implementation called \NAME.  Our simulations confirmed that
\processes quickly reach consistent partitioning together, and that
the generated traffic remains locked down to partitions.  We also
briefly discussed \NAMEC, an improvement to contain the traffic in relevant
autonomous systems.
%% correctness of our protocol as well as the confinement of
%% the traffic only where it is needed.

%% As future work, we plan to leverage the hierarchical properties of
%% interconnected autonomous systems to further limit the propagation of
%% indexes within interested systems only. We expect that such an
%% improvement would greatly benefit the autonomous systems, and
%% particularly those hosting a large number of contents.

%% Systems hosting a large number of contents would greatly benefit from
%% this improvement.

%% As future work, we plan to enable lazy retrieval of partitions when a
%% large number of contents is involved. Indeed, using \NAME, every
%% \process partakes in the propagation of all indexes which raises
%% scalability issues. We would like to leverage the hierarchical
%% properties of
%% % interconnected
%% autonomous systems to limit the
%% propagation of such indexes only to interested systems.

%
%% As ongoing and future work, we are currently studying how it might be
%% possible to limit the scope of a partition to one autonomous system at
%% most. Indeed, the gain of propagating changes to \processes belonging to
%% another autonomous system might be negligible as the latency penalty
%% is mainly due to inter autonomous systems' links (\eg trans-continental
%% links).  Exactly, we are investigating how such a change can be
%% formalized and its impact on the current protocol.
%
%autonomous systems. Indeed, \processes could leverage the hierarchical
%properties of the Internet topology~\cite{nur2018geography} to avoid
%flooding the whole systems with control information about all
%partitions.
%\TODO{should we highlight this at the end of the experiment}
%

At mid-term, we plan to evaluate our proposal within a concrete
storage system such as \underline{I}nter\underline{P}lanetary
\underline{F}ile \underline{S}ystem
(IPFS)~\cite{henningsen2020mapping}. This would allow us to assess the
relevance of the \NAME approach in a real system subject to high
dynamicity, and compare it against its current DHT-based indexing
system that does not include distance in its operation. More generally,
we claim \NAME (or \NAMEC) can constitute a novel building block for
geo-distributed services.  Among the use-cases we have identified,
\NAME can be used to complement content delivery infrastructures by
efficiently sharing between \process attached to the same CDN server
the locatiosn of web objects that have been already downloaded. This
would allow the containment of web traffic even lower in our networks,
and so reduce the overall traffic footprint.

%% Such an integration would enable us to to perform additional
%% experiments and observe the gain with respect to the DHT-based
%% approach that is currently used.

%% In addition to this action, it would make sense to evaluate
%% the use of \NAME within a concrete storage systems, such as
%% \underline{I}nter\underline{P}lanetary \underline{F}ile
%% \underline{S}ystem (IPFS)~\cite{henningsen2020mapping}.  Such an
%% integration would enable us to to perform additional experiments and
%% observe the gain with respect to the DHT-based approach that is
%% currently used.

%
%% More generally, it can be used in other services in order to share information while mitigating the traffic.
%% MWhile we use the distance to the closest copy in this paper, It is
%% notewothy that other scopes can be envisioned according to the
%% use-cases.  For instance, the location of a city in order to broadcast
%% message to all nodes in the city.  This requires nodes to store and
%% maintain their city location in their local state. Nodes stop
%% delivering and forwarding their received messages when they come from
%% a different city. Such a predicate can be complexified in order to
%% broadcast messages to all nodes in the city plus neighboring
%% cities. This requires nodes to overload forwarded messages the first
%% time they reach another city. The predicate checks if messages already
%% reached two distinct cities before delivery.  The predicate can even
%% be piggypack within the message and so be different for each
%% message.\TODO{Complete the use-cases}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "../paper"
%%% ispell-local-dictionary: "english"
%%% End: