### 2022-06-01 rebuttal2.txt
rebutall draft

We thank the three reviewers for their thorough reading and
interesting comments.

* About motivation and practical application: (TODO) Add reference to Bastien's paper
A user from Europe wants to download a video indexed only in
Australia despite a neighbor
European having a replica (because he was interested in the video as
well). Still, the former originates a round-trip to Australia only to
discover that it wasted time looking for the location of the video
that was next to him.  DHTs also have this pitfall where nodes knowing
the video locations might be running in Australia. Our approach
proposes a trade-off that allows close node to answer location
requests, even in dynamic networks where nodes can quickly reconfigure
to meet the changes in a peer-to-peer fashion.
** Action: We use such an example throughout the paper, complementing
the figure descriptions with real-world usage and location.
** Side note: Out of the scope of networking, our approach also
constitutes an efficient way to perform a multi-destinations shortest
path in dynamic graphs.

* About failure detection: We agree with Reviewer 2 that implementing
link failure detection is often done using heartbeats. Our approach is
built on top of such links and benefits, as many other running
protocols, from such failure detection. Since this is a shared overhead,
we do not account it as being part of our protocol.
** Action: Clarify this prerequisite and its implication on the overhead.

* About crash recovery: As of now, crashes may have immediate impact
on neighboring nodes depending on their crash detection
speed. Therefore, a node must always restart anew, thus retrieving and
propagating relevant information again.
** Side note: There exists a trade-off between convergence speed and
and generated traffic when buffering incoming messages and only
propagating most relevant ones. We kept this out of scope for the
sake of simplicity.

* About the experiments: PeerSim, despite being old, remains a
flexible framework to run large scale simulations. It fits our goal of
isolating the behavior of our protocol under desired circumstances.
Nevertheless, we agree with Reviewer 1 that more realistic
configurations are of interest. We left this as future work, where we
would like to replace the DHT of IPFS by our protocol, and deploy it
on a real distributed infrastructure.

** About Figure 7 link: We agree with Reviewer 2 that a single
intercontinental link is not realistic, however this allows us
to confirm and highlight the very important property that our
protocol works even in a network with physical partitions. // (TODO) more ?
** About Figure 7 time: We use simulated time as x-axis to highlight that
the convergence speed depends on the simulated latency between links.
** Action: Use both cycle and simulated time as x-axis label.
** About simulated time and real network conditions: we only simulated 
the link latency using event-based PeerSim, i.e., the number of cycles
(simulated milliseconds to help human understanding) between the
sending of a message and its receipt by a neighbor. Other network conditions
such as hardware limitations, packet loss, would indeed increase convergence
time, but we expect it to remain in the same order of magnitude when not catastrophic.
// (TODO) rewrite this…

** Comparison with state-of-the-art: As stated in Related Work,
state-of-the-art approaches have shortcomings that disqualify them as
contender in our context: Centralized and DHT approaches have slow
response time; vector-based routing and replicated stores flood the
whole network; timeout-based, random walks, and spanning forests are
cycle-based which never-endlessly floods the network.

// 582 words [2022-06-01 Mer] 11:58