### 2022-06-02 rebuttal2.md
985 words

We thank reviewers for their thorough reading and interesting
comments. We hope that our response will clarify their interrogations.

# Reviewer 1

## Motivation and practical application:
The starting point of our study is about minimizing response time of
locating content in geo-distributed infrastructures. Each time someone
looks for a particular content, it must start by locating it in the
infrastructure: a first node is reached to determine its location and
then, the content is retrieved from a second node [1]. "Retrieving a
content location may actually take more time than retrieving the
content itself".

For instance, let us consider Alice from Vienna who wants to download
a video. She first needs to know where to retrieve it from, and so
asks a content location service. If the latter is centralized and
hosted far away in Australia, the round-trip time to interact with the
location service is detrimental, especially when the video is actually
stored close to Vienna.

One might think that distributing the indexing service using DHTs
would be sufficient to avoid this pitfall. However, DHTs do not
collocate content and location services. Therefore, despite having a
DHT node in Vienna, the one that actually knows where to find the
video Alice is looking for is still in Australia. Hence, the
interaction with the location service is still time-consuming.

Using our approach, each node in charge of maintaining content
locations knows at any time the closest location for a particular
content. This allows Alice to contact its close Vienna location
service to then download the video from her closest location, thus
improving response time.

Our approach is not limited to content indexing, and generalizes to
any replicated service. On a side note, it even constitutes an
efficient way to perform multi-destinations shortest path in dynamic
graphs.

### Action:
Slightly revise the introduction to instantiate this Vienna example
and complement the description of figures to further help the reader.

## Simulations vs emulations vs in-vivo:
PeerSim, despite being old, remains a simple and flexible framework to
run large scale simulations in comparison to Splay, Mininet etc.  It
fits our goal of isolating the behavior of our protocol under desired
circumstances.

We focus our experiments on link latency. Other network conditions
such as hardware limitations and heterogeneity, packet loss, link
congestion, would indeed increase convergence time, but we expect it
to remain in the same order of magnitude when not catastrophic.

Nevertheless, we agree that more realistic configurations are of
interest. We leave this as future work, where we would replace the DHT
of IPFS by AS-cast, and deploy it on a real distributed
infrastructure.

## Comparison against state-of-the-art:
While previous works proposed such comparisons [1], we voluntarily
discard them as they have well-identified shortcomings: Centralized
and DHT approaches do not collocate content and location services,
leading to slow response time; vector-based routing and replicated
stores flood the whole network at each event; timeout-based, random
walks, and spanning forests are cycle-based which never-endlessly
floods (part of) the network.

## Figure 7 time:
The time reported on Figure 7 corresponds to the simulated one. A
cycle is a simulated millisecond.

### Action:
Underline in the legend that every reported time is a simulated one.

# Reviewer 2

## Centralized and DHT
Please, look at the first anwser provided to Reviewer 1.

## Node failures:
Algorithm 3 is designed to handle changes in the infrastructure.
"Removing a node is equivalent to removing all its edges". Therefore,
"What happens if this direct parent fails?": Algorithm 3, line 4
states that this event is equivalent to a possible deletion. The
detecting node must propagate the changes to its neighbors
accordingly.

Regarding the detection of failures, we indeed assume a low level
mechanism, such as heartbeats.  AS-cast is built on top of such a
mechanism, and as many other running protocols, benefits from such
failure detection.

### Action:
We clarify this requirement and its implication on the overhead.

## Scenarios 3e and 5e
Figure 3 illustrates the challenge of maintaining consistent
information.  Figure 5 demonstrates that Node C cannot distinguish the
two scenarios.  The node must assume the worst possible outcome and
propagate delete messages in both cases, that would be true in Figure
3f, but false positives in Figure 5f. AS-cast guarantees that every
node converge to its proper partition afterwards.

## First evaluation practicability
The topology is indeed arbitrary as we do not find any real
geo-distributed topology reaching such a scale of 10K nodes.  We
thought of instantiating multiple times the GEANT topology but
interconnecting them would also be arbitrary.

## About worst-case scenario
Let us consider a chained set of Z nodes [A-Z] and the sequential
creation of partitions, the number of messages in the worst case is:
25×2 + 24×2+ … + 1×2 = O(2×Z^2).

When creating the same partitions in parallel, the number of messages
is only: 2×25.

## Size of packets
We agree that transferred bytes would be of interest, especially to
compare with other approaches in more realistic settings as we plan in
future work.

## Single link between continents
We agree that this scenario is unrealistic; and also that partitions
may span multiple continents. AS-cast would work in such scenario.
This experiment aim to demonstrate that AS-cast behaves correctly in
case of physical partitions. We find this result most interesting
since decentralized protocols often fail to do so.

# Reviewer 3

## Content indexing formal problem 
We chose to abstract the content indexing challenge as a more general
partitioning problem and so formalize it this way (see also first
answer to R1-R2).

## Scope predicate
This is indeed an interesting point. We thank to reviewer for this
valuable remark and let it as future work.

## Black-and-white
### Action:
We agree and make colored figures black-and-white friendly.

## Crash recovery
Crashes may have immediate impact on neighboring nodes.  Therefore, a
node always restart anew, retrieving and propagating relevant
information again.

[1] https://hal.archives-ouvertes.fr/hal-02190125/file/tnsm.pdf