### 2022-06-03 rebuttal2.md
809 words

We thank reviewers for their thorough reading and interesting
comments. We hope that our response will clarify their interrogations.

# Reviewer 1

## Motivation and applications:
Our starting point is about minimizing response time of locating
content in geo-distributed infrastructures. When looking for a content,
a node is reached to determine its location and then, the content is
retrieved from another node [1]. __"Retrieving a content location may
actually take more time than retrieving the content itself"__.

Let us consider Alice from Vienna who wants to download a video. She
must know where to retrieve it from, therefore she asks a content
location service. If the latter is centralized and hosted faraway in
Australia, the round-trip to the service is detrimental, especially
when the video is actually stored close to Vienna.

Distributing indexing services using DHTs is insufficient to avoid
this pitfall: DHTs do not collocate content and location
services. Therefore, despite having a DHT node in Vienna, the one that
actually knows where to find the video Alice is looking for is still
in Australia. Interacting with the location service remains
time-consuming.

AS-cast improves quality of service since each node in charge of
maintaining content locations always knows the closest location for a
particular content.  Alice contacts her Vienna location service then
downloads the video from her closest host.

AS-cast is not limited to content indexing, and generalizes well to
any replicated service. It even constitutes an efficient way to
perform multi-destinations shortest path in dynamic graphs.

**Action:** Slightly revise the introduction to instantiate this
Vienna example and complement figure descriptions.

## Simulations vs emulations vs in-vivo:
PeerSim remains a simple and flexible framework to run large scale
simulations in comparison to Splay, Mininet etc.  It fits our goal of
isolating the behavior of our protocol under desired circumstances.

We focus our experiments on link latency. Other network conditions
such as hardware limitations/heterogeneity, packet loss, link
congestion, would indeed increase convergence time, but we expect an
identical order of magnitude when not catastrophic.

Nevertheless, we agree that more realistic configurations are of
interest. We leave this as future work: replace the DHT of IPFS by
AS-cast, and deploy it on a real distributed infrastructure.

## State-of-the-art comparison:
Previous works proposed such comparisons [1]. We discard them as they
have well-identified shortcomings: Centralized and DHT approaches do
not collocate content and location services, leading to slow response
time; vector-based routing and replicated stores flood the whole
network at each event; timeout-based, random walks, and spanning
forests are cycle-based which never-endlessly floods (parts of) the
network.

## Figure 7 time:
The time reported on Figure 7 corresponds to the simulated one. A
cycle is a simulated millisecond.

# Reviewer 2

## Centralized and DHT
Please, read the first answer provided to Reviewer 1.

## Node failures:
Algorithm 3 handles infrastructure changes.  __"Removing a node is
equivalent to removing all its edges"__.  Therefore, "What happens if
this direct parent fails?": Algorithm 3, line 4 triggers a _possible_
source deletion. The detecting node propagates the changes to its
neighbors accordingly.

To detect failures, we indeed assume a low level mechanism, such as
heartbeats. Since many other running protocols benefit from provided
failure detection, we do not consider it as an overhead of our
protocol.

**Action:** Clarify requirements and overhead implications.

## Scenarios 3e and 5e
Figure 3 illustrates the challenge of maintaining consistent
information.  Figure 5 highlights the challenge that Node C cannot
distinguish the two scenarios.  The node assumes the worst and
propagates delete messages in both cases, rightfully in Figure 3f,
wrongfully in Figure 5f. Afterwards, AS-cast ensures that every node
converges to its correct partition.

## First evaluation practicability
The topology is indeed arbitrary as we do not find any real
geo-distributed topology reaching 10K nodes.  Instantiating GEANT
multiple times would remain arbitrary.

## Worst-case scenario
Let us consider [A..Z] chained nodes, and a sequential creation of
partitions, the number of messages is: 25×2 + 24×2 +…+ 1×2 = O(2×Z^2).
Whereas in parallel, the number of messages is only: 2×25.

## Packets size
We agree that transferred bytes would be interesting, especially to
compare with other approaches in realistic settings as we plan in
future work.

## Intercontinental link
We agree that this scenario is unrealistic.  Nevertheless, this
experiment aims to demonstrate that AS-cast behaves correctly in case
of physical partitions. We find this result most interesting since
decentralized protocols often fail in such context.

# Reviewer 3

## Content indexing problem 
We chose to abstract the content indexing challenge as a more general
partitioning problem and formalize it this way.

## Scope predicate
Any predicate would work: the scope of an always true predicate would
trivially be "all nodes".

## Black-and-white
**Action:** Make figures black-and-white friendly.

## Crash recovery
Crashes may impact neighboring nodes immediately.  Therefore,
nodes always restart anew, retrieving and propagating relevant
information again.

[1] https://hal.archives-ouvertes.fr/hal-02190125/file/tnsm.pdf