### 2022-06-28 input/appendices.tex
major changes everywhere

\appendix

\section{Table of notations}
\label{appendix:notations}
\input{input/tablenotations.tex}

%% Table~\ref{table:notations} summarizes the notations used throughout
%% this paper. For the sake of clarity, we divide them into three
%% categories: network, \process, and message.

% \section{Appendix: Proofs of Theorem~\ref{theo:fb} and Theorem~\ref{theo:fbfsdcp}}
% \label{appendix:fb}

%%\section{Proof of Theorem~\ref{theo:fbfsdcp}}
%% \label{appendix:fbfsdcp}