### 2022-06-28 input/figdel.tex
reworkd figures so they dont overlap, added a figure about naive deletes

\begin{figure}[t]
  \newcommand{\SCALE}{0.8}

  \newcommand\X{42pt}
  \newcommand\Y{-40pt}

  \thickmuskip=0mu
  \medmuskip=0mu
  \thinmuskip=0mu

  \newcommand\INVISIBLESPACING{1.6}
  \newcommand\ADD{\alpha}
  \newcommand\DEL{\delta}

  \newcommand{\OACK}{0.5} %% opacity of acknowledgement alpha messages

  \newcommand{\LEFT}{\triangleleft}
  \newcommand{\RIGHT}{\triangleright}

  \begin{center}
    \subfloat[Part A][\label{fig:delA}$a$ deletes its partition. It
      notifies all \processes that belong to its
      partition.]{\input{input/figdelA.tex}}
    \hspace{3pt}
    \subfloat[Part B][\label{fig:delB}$\delta$ stops when it encounters
      another partition. $b$ answers with its partition.]{\input{input/figdelB.tex}}
    \hspace{3pt}
    \subfloat[Part C][\label{fig:delC}$a$ receives, delivers, and forwards the message $\alpha_d^3$.
      Every node belongs to $P_d$.]{\input{input/figdelC.tex}}
  \end{center}
  \caption{\label{fig:del}Efficient removal of a partition using
    scoped broadast.} % $a$ eventually acknowledges that it belongs to $P_d$ with $\alpha_d^3$ echoing.}
\end{figure}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "../paper"
%%% ispell-local-dictionary: "english"
%%% End: