### 2022-10-19 input/figproblem.tex

\begin{figure*}
  \begin{center}
    \subfloat[Part A][\label{fig:problemA}Both nodes $a$ and $d$ create a
      new partition.  $w_{ab} = 2$; $w_{bc} = w_{bd} = 1$; $w_{cd} =
      3$.]{\input{input/figproblemA.tex}}
    \hspace{0pt}
    \subfloat[Part B][\label{fig:problemB}Both $a$ and $d$ delete
      their partition. $b$ delivers and forwards $\alpha_a$ while
      $c$ delivers and forwards both $\alpha_d$ and $\delta_d$.]
             {\input{input/figproblemB.tex}}
    \hspace{0pt}
    \subfloat[Part C][\label{fig:problemC}$b$
      already blocks the propagation of $\delta_a$ despite it
      propagated $\alpha_a$ before; and $c$ does not deliver
      $\alpha_d$, for it already deleted it
      before.]{\input{input/figproblemC.tex}}
    \hspace{0pt}
    \subfloat[Part D][\label{fig:problemD}\Processes
      discard remaining messages. Since $b$ cannot deliver
      $\delta_a$, $c$ and $d$ remain in the deleted $P_a$.]
             {\input{input/figproblemD.tex}}
  \end{center}
  \caption{\label{fig:problem}Naive propagation of $\alpha$ and
    $\delta$ messages is insufficient to guarantee consistent partitioning.}
\end{figure*}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "../paper"
%%% ispell-local-dictionary: "english"
%%% End: