### 2022-10-19 input/prooftheoremfb.tex

Whenever a \process~$s$ becomes a source, it broadcasts hence
delivers its own message $d_s(\alpha_{s}^{|\pi_{ss}|})$. Whatever
its set of received messages, it acknowledges that it belongs to its
own partition $P_s$ since $\alpha_{s}^{|\pi_{ss}|} = \min D_s$ and
it remains forever since $|\pi_{ss}|$ is its greatest lower bound.
\noindent Such a source forwards its notification to its neighbors.
Every neighbor eventually receives its notification since
communication links are reliable. Whatever the order of received
messages $R_x$ at neighboring \process~$x$, total order ensures that
it delivers notifications $\alpha_s^{d}$ when
$d= |\pi_{ss}| + w_{sx} = \min |\pi_{s'x}|$ where $\min |\pi_{s'x}|$
is the lightest weight of $x$ to any source $s'$, $s'$ being $s$ in
this case.
\noindent Among these neighbors, at least those that fulfill the
latest condition forward their respective notification.  By
transitivity, the message originating from $s$ reaches all
\processes that belong to $P_s$ at least through their respective
lightest path: $\small\smash{\forall y \in V, s, s' \in S: \min
  |\pi_{s y}| < \min |\pi_{s' y}| \implies \eventually
  d_{y}(\alpha_{s}^{\min |\pi_{sy}|})}$. When the system becomes
quiescent, \ie no \process becomes source anymore, every \process
eventually acknowledges the partition it belongs to, \ie \processes
eventually reach consistent partitioning together. In addition, the
protocol terminates: a \process never delivers hence forwards a
message after it received, delivered, and forwarded the message of
its closest source from its lightest path.