### 2022-10-19 input/prooftheoremfbfsdcp.tex

Detection triggers forwarding of staleness downstream which
completes the case study of Lemma~\ref{lem:detector} by ensuring
that, when a source broadcasts its staleness notification, all
\processes that belong to its partition eventually deliver such a
staleness notification. All downstream bordering \processes also
eventually receive such a staleness notification and echo back their
best delivered message. This triggers another competition as in
Theorem~\ref{theo:fb} for the \processes that delivered the
staleness notification.